---
title: "Homework #2"
author: "Kelly Paek"
date: "2024-09-25"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Question #1
```{r Question 1}
data("diamonds")
#setting the x & y axis and variable for color in ggplot
ggplot(data = diamonds, aes(x = carat, y = price, color = color)) + 
  #setting the plot type to point plots
  geom_point(alpha = 0.3) +
  #renaming labels on x and y axis and color legend
  labs(x = "Carat", y = "Sale Price ($)", color = "Color Grade") +
  #setting the theme of the plot to minimal
  theme_minimal() +
  #changing the scale of the x axis to the log base of 2 to replicate the given graph.
  scale_x_continuous(trans='log2')

```

## Question #2 - Part A
```{r Question 2A}
# install.packages("babynames")
library(babynames)
data("babynames")

#subsetting so that df only contains desired names
baby_subset <- subset(babynames, subset = name %in% c("Ryan", "Jeff", "Shonda", 
                                                    "Jonathan", "Collin", 
                                                    "Nathan", "William"))

ggplot(baby_subset, aes(x = year, y = n, color = name)) + geom_line() 
```
The graph above is messy because every name occurs twice for each year (for male and female) in the dataset, and for every name in every year, ggplot is plotting 2 values at once, creating this zigzag line graph. Also, this is not a great way to represent this dataset because there are way more 'William's in the dataset compared to other names, which makes it hard to see the plots for other names. To fix this issue, we can facet wrap to create individual plots for each name, and also plotting separate lines for sex. 


## Question #2 - Part B
```{r Question 2B}
#plots frequency of each name over time, and distinguishes frequency by sex
ggplot(baby_subset, aes(x = year, y = n, color = sex)) + 
  geom_line() +
  #wrapping the plots by name so there is a plot for each name
  facet_wrap(~name) +
  #renaming labels to make plot neater
  labs(x = "Year", y = "Frequency per year", color = "Sex") +
  #ressigning colors of line plots and legend labels
  scale_colour_manual(values = c("hotpink3", "royalblue3"), 
                      labels = c('Female', 'Male')) +
  theme_minimal()

```
*help received from mentor Siho*

## Question #3 - Part A
```{r Question 3A}
data("economics")

#subsetting so that df only contains columns needed for 3B
new_economics <- subset(economics, select = -c(pce, pop, unemploy))

head(new_economics, 3)
```

## Question #3 - Part B
```{r Question 3B}

#making the df longer so type of outcome 'psavert' or 'uempmed' is set as a
#categorical variable
b_economics <- pivot_longer(new_economics, 
                           cols = c("psavert", "uempmed"), 
                           names_to = "Outcome", 
                           values_to = "Values")
ggplot() +
  #plotting to display 'psavert' and 'uempmed' according to time
  geom_line(data = b_economics, 
            aes(x = date, y = Values, color = Outcome)) +
  labs(x = "Age", y = "Height") +
  theme_minimal()

```
These variables appear to be related, because if the uempmed(Median Weeks Unemployed) increases, psavert(Personal saving rate) increases, and the vise versa.

*Collaborated with Debanjali*

## Question #4
```{r Question 4}
library(datasets)
data("Loblolly") 


avgLoblolly <- Loblolly %>%
  #grouping the df by age
  group_by(age) %>%
  #creating a new column that contains the mean height for each age
  mutate(avgheight = mean(height)) %>% 
  #selecting only the age and average height from the df
  select(age, avgheight)

ggplot() +
  #plotting the new df from above (avgheight vs height) as a red line
  geom_line(data = avgLoblolly, 
            aes(x = age, y = avgheight), 
            color = "red",  
            lwd = 2) +
  #plotting the original df (height of each tree), with each line representing
  #the height of each tree on another layer
  geom_line(data = Loblolly, 
            aes(x = age, y = height, group = Seed), 
            alpha = 0.5) +
  #reassigning labels
  labs(x = "Age", y = "Height") +
  #changed theme to recreate given plot
  theme_bw()

```
*Collaborated with Debanjali and Youssef*


## Question #5 - Part A
```{r Question 5A}
library(Lahman)
data("Teams")

#create a new df
top_teams <- Teams %>% 
  #create a new column with newly calculated total bases
  mutate(TB = ((H - X2B - X3B - HR)*1 + X2B*2 + X3B*3 + HR*4)) %>%
  #create a new column for SLG
  mutate(SLG = TB/AB) %>%
  #selecting only the desired columns
  select(yearID, teamID, SLG) %>% 
  ### Modification: filtered to only select top teams since 1969
  filter(yearID >= 1969) %>% 
  #arranging by the descending order for SLG to select top 10 teams
  arrange(desc(SLG)) %>% 
  #select top 10 teams
  head(., 10)

top_teams
```

## Question #5 - Part B
```{r Question 5B}
data("Batting")

top_player <- Batting %>%
  #choose player from Batting from teams from Part A
  semi_join(top_teams, by = c("yearID", "teamID")) %>%
  #group df by yearID and teamID
  group_by(yearID, teamID) %>%
  #pick the player with the highest HR
  top_n(1, HR) %>%
  #select desired columns
  select(yearID, teamID, playerID, HR)

top_player
```


