---
title: "Homework #3"
author: "Kelly Paek"
date: "2024-09-24"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)
library(dplyr)
```

## Question #1
```{r Question 1}
polls <- read.csv("https://remiller1450.github.io/data/polls2016.csv")

# make dataset longer so that each candidate (Trump and Clinton) can act as a 
# catergorical variable
new_polls <- polls %>% pivot_longer(cols = c("Trump..R.", "Clinton..D."),
                                    names_to = "Candidate",
                                    values_to = "Support") %>%
  # only select necessary columns
  select(., c("Poll", "Candidate", "Support"))

# creating the plot
ggplot(data = new_polls, aes(x = Candidate, y = Support, fill = Candidate)) +
  # ensure that pre-calculated support percentages are used as the y
  geom_bar(stat = "identity") +
  # facet by each poll
  facet_wrap(~Poll) +
  # assign correct colors to the fill according to the candidates
  scale_fill_manual(values = c("blue", "red"),
                    labels = c("Clinton..D.", "Trump..R.")) +
  # add plot title and remove x title
  labs(title = "Candidate Support by Poll", x = NULL) +
  theme_minimal()

```



## Question #2

Long data is data with less columns with and more rows. Each observation is broken down to multiple rows to represent the different aspects of the observation. They are harder for us, humans, to understand the data, but easier to manipulate with R (depending on the work that needs to be done with the dataset). They are used for analyses requiring multiple observations per subject.

Wide data, on the other hand, has a lot more columns. Each variable has its own column and each row represents a subject; multiple measurements/observations are spread across multiple columns. They are easier to read and understand with the human eye.

To convert long to wide format, we use pivot_wider(df, id_cols, names_from, values_from). 'id_cols' would be the variable that would be given its own row. 'names_from' would be the column where all the column names would be coming from, and 'values_from' would be where the actual data would be coming from.

To convert wide to long format, we use pivot_longer(df, cols, names_to, values_to). 'cols' would be the what would be included in the long, single list of values. 'names_to' would be the name of column that would contain all the column names of the previous version of df. 'values_to' would be the name of the column where all the values would be saved in.

*referred to our Lab 6 webpage for help*

## Question #3
```{r Question 3}
icu <- read.csv("https://remiller1450.github.io/data/ICUAdmissions.csv")

# Filter for infection in ICU visits and Previous == 0
new_icu <- icu %>%
  filter(Previous == 0) %>%
  # group by sex so we get z score by sex
  group_by(Sex) %>% 
  # Create and calculate new column for age-specific Z-scores
  mutate(age_spec_Z_score = 
           (Age - mean(Age, na.rm = TRUE)) / sd(Age, na.rm = TRUE))

# plotting a density chart of sex specific Z-scores for each sex
ggplot(new_icu, aes(x = age_spec_Z_score, y = HeartRate, fill = as.factor(Sex))) +
  # modifying labels for legend and changing color of plots
  scale_fill_manual(labels = c("Male", "Female"),
                    values = c("black", "red")) +
  scale_color_manual(values = c("black", "red")) +
  geom_histogram(mapping = aes(y = after_stat(density), color = as.factor(Sex)), 
                 position = "identity", 
                 alpha = 0.5,
                 bins = 30,
                 na.rm = TRUE) +
  # change title and axes labels according to the given plot
  labs(title = "HR Density by Sex", x = "Age Z-score", y = NULL, fill = "Sex") +
  # change theme to recreate given plot
  theme(panel.background = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  guides(color = "none")

```

```{r}
drives = read.csv("https://remiller1450.github.io/data/drives.csv")
```

*Collaborated with Cassidy, Debanjali, and mentor Siho*
